---
title: "LSA methods comparison R"
author: Anton Antonov
date: 2022-02-27
output: html_notebook
---


<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

-------

# Get Python mandalas collection

## Setup

```{r setup, echo=FALSE, include=FALSE}
library(Matrix)
library(magrittr)
library(purrr)
library(LSAMon)
library(SparseMatrixRecommender)

library(reticulate)
#reticulate::py_config()

source("../R/UtilityFunctions.R")
```

```{python}
# "Standard" packages
import pandas
import numpy
import random
import io
import time

# SSparseMatrix, SMR, and LSA packages
from SSparseMatrix import *
from SparseMatrixRecommender import *
from LatentSemanticAnalyzer import LatentSemanticAnalyzer

# Plotting packages
import matplotlib
import matplotlib.pyplot as plt
from mpl_toolkits.axes_grid1 import ImageGrid

# Image processing package(s)
from PIL import Image, ImageOps

# Random mandalas packages
from RandomMandala import random_mandala, figure_to_image
```

## Mandala collection

```{python}
# A list to accumulate random mandala images
mandala_images = []

# Generation loop
random.seed(443)
tstart = time.time()
for i in range(512):
    
    # Generate one random mandala figure
    fig2 = random_mandala(n_rows=None,
                          n_columns=None,
                          radius=[8, 6, 3],
                          rotational_symmetry_order=6,
                          symmetric_seed=True,
                          connecting_function='bezier_fill',
                          face_color="0.", 
                          alpha = 1.0)
    fig2.tight_layout()
    
    # Convert the figure into an image and add it to the list
    mandala_images = mandala_images + [figure_to_image(fig2)]
    
    # Close figure to save memory
    plt.close(fig2)

# Invert image colors    
mandala_images1 = [ImageOps.invert(img) for img in mandala_images]

# Binarize images
mandala_images2 = [im.convert('1') for im in mandala_images1]

# Resize images
width, height = mandala_images2[0].size
print([width, height])
ratio = height / width
new_width = 200
mandala_images3 = [img.resize((new_width, round(new_width * ratio)), Image.ANTIALIAS) for img in mandala_images2]
print("Process time: " + str(time.time()-tstart))
```

## Get mandalas in R

```{python}
mandala_arrays = [numpy.asarray(x, dtype="int32") for x in mandala_images3]
```


```{r}
length(py$mandala_arrays)
```

```{r}
# This plots the 512 images. Using {r include=FALSE} produces some strange behaviour.
# Hence we just the arrays below.

# system.time({
#   mandalas3 <- purrr::map(py$mandala_arrays, function(x) image(x, axes = FALSE, col = gray( 0:255 / 255 ) ))
# })
```


------

# Process random mandalas collection

Get the mandala arrays:

```{r}
mandalaImageArrays <- purrr::map(py$mandala_arrays, t)
```

Show one of the random mandalas:

```{r, fig.width=2, fig.height=2}
image(mandalaImageArrays[[12]], axes = FALSE, col = gray(0:255/255) )
```
Show an array of inverted random mandalas:

```{r}
set.seed(332)
ParImageGrid(sample(mandalaImageArrays, 12), mfrow = c(3,4))
```

Convert each image array into a flatten array:

```{r}
mandalaArrays <- purrr::map(mandalaImageArrays, as.numeric)
length(mandalaArrays)
```

Make a matrix of flattened mandalas:

```{r}
mandalaMat <- matrix(do.call(c, mandalaArrays), ncol = length(mandalaArrays[[1]]), byrow = TRUE)
dim(mandalaMat)
```


Make a sparse matrix of flattened mandalas:

```{r}
mandalaSMat <- as(mandalaMat, "sparseMatrix")
mandalaSMat <- as(mandalaSMat, "dgCMatrix")
mandalaSMat <- drop0(mandalaSMat)
dim(mandalaSMat)
length(mandalaSMat@x)/length(mandalaSMat)
```

-----

# LSAMon object creation

Se the number of topics:

```{r}
numberOfTopics <- 40
```

Create the `LSAMon` object:

```{r}
system.time(
  lsaObj <- 
    LSAMonUnit() %>% 
    LSAMonSetDocumentTermMatrix(mandalaSMat) %>% 
    LSAMonApplyTermWeightFunctions(globalWeightFunction = "None", localWeightFunction = "None", normalizerFunction = "None")
)
```

## SVD

Here we extract image topics using Singular Value Decomposition (SVD):

```{r}
system.time(
  lsaSVDObj <- 
    lsaObj %>%
    LSAMonExtractTopics(numberOfTopics = numberOfTopics, method = "SVD", maxSteps = 120, minNumberOfDocumentsPerTerm = 0)
)
```

```{r}
svdH <- lsaSVDObj %>% LSAMonNormalizeMatrixProduct(normalizeLeftQ = FALSE) %>% LSAMonTakeH

dim(svdH)
length(svdH@x)/length(svdH)
```


## NNMF

Here we extract image topics using Non-Negative Matrix Factorization (NNMF):

```{r}
system.time(
  lsaNNMFObj <- 
    lsaObj %>%
    LSAMonExtractTopics(numberOfTopics = numberOfTopics, method = "NNMF", maxSteps = 16, minNumberOfDocumentsPerTerm = 0)
)
```

```{r}
nnmfH <- lsaNNMFObj %>% LSAMonNormalizeMatrixProduct(normalizeLeftQ = FALSE) %>% LSAMonTakeH

dim(nnmfH)
length(nnmfH@x)/length(nnmfH)
```

-------

# Show topics interpretation

## SVD

Show the importance coefficients of the topics (if SVD was used the plot would show the singular values):

```{r}
smat <- lsaSVDObj %>% LSAMonNormalizeMatrixProduct(normalizeLeftQ = TRUE) %>% LSAMonTakeH
lsNorms <- sqrt(rowSums(smat * smat))
barplot(lsNorms)
```


```{r SVD-basis-images}
lsBasisImages <- 
  purrr::map(1:nrow(svdH), function(i) { 
    matrix(svdH[i,], nrow = nrow(mandalaImageArrays[[1]]), byrow =F)
  })
lsBasisImagesLabels <- 1:length(lsBasisImages)

par( mfrow = c(5,8), mai = c(0,0,0,0))
for(i in 1:length(lsBasisImages)){
  image( lsBasisImages[[i]], axes = FALSE, col = gray( 0:255 / 255 ) )
  text( 0.2, 0, lsBasisImagesLabels[[i]], cex = 1.2, col = 2, pos = c(3,4))
}
```

## NNMF

Show the importance coefficients of the topics (if SVD was used the plot would show the singular values):

```{r}
smat <- lsaNNMFObj %>% LSAMonNormalizeMatrixProduct(normalizeLeftQ = TRUE) %>% LSAMonTakeH
lsNorms <- sqrt(rowSums(smat * smat))
barplot(lsNorms)
```

Drop "insignificant" pixels:

```{r clip-NNMF-basis-values}
summary(nnmfH@x)
nnmfH2 <- nnmfH
if( TRUE ) {
  nnmfH2@x[nnmfH@x < 1.5e-2] <- 0
  nnmfH2 <- drop0(nnmfH2)
}
```

Show the interpretation of the extracted image topics:

```{r NNMF-basis-images}
lsBasisImages <- 
  purrr::map(1:nrow(nnmfH), function(i) { 
    matrix(nnmfH2[i,], nrow = nrow(mandalaImageArrays[[1]]), byrow =F)
  })
lsBasisImagesLabels <- 1:length(lsBasisImages)

par( mfrow = c(5,8), mai = c(0,0,0,0))
for(i in 1:length(lsBasisImages)){
  image( lsBasisImages[[i]], axes = FALSE, col = gray( 0:255 / 255 ) )
  text( 0.2, 0, lsBasisImagesLabels[[i]], cex = 1.2, col = 2, pos = c(3,4))
}
```

-------

# Approximation

Get a new, unseen random mandala:

```{python}
random.seed(89)


# Consider using radius = [8, 6, 3]
fig = random_mandala(radius=[8, 6, 3],
                     rotational_symmetry_order=6,
                     symmetric_seed=True,
                     connecting_function='bezier_fill',
                     face_color="0.", 
                     alpha = 1.0)
                      
# Convert the mandala figure to image
testMandala = figure_to_image(fig)

# Resize image
testMandala1 = testMandala.resize((new_width, round(new_width * ratio)), Image.ANTIALIAS)

# Invert image colors    
testMandala2 = ImageOps.invert(testMandala)

# Binarize image
testMandala3 = testMandala2.convert('1')

# Resize image
testMandala4 = testMandala3.resize((new_width, round(new_width * ratio)), Image.ANTIALIAS)

# The image
testMandala4
```

Plot the image:

```{python}
plt.imshow(testMandala4)
plt.axis('off')
plt.show()
```
Get sparse vector corresponding to the generated mandala image in R:

```{python}
testMandalaArray = numpy.asarray(testMandala4, dtype="int32")
```

```{r}
testMandalaArray <- py$testMandalaArray
dim(testMandalaArray)
```

```{r}
testMandalaSMat <- sparseMatrix( x = as.numeric(t(testMandalaArray)), i = rep_len(1,length(testMandalaArray)), j= 1:length(testMandalaArray))
testMandalaSMat <- as(testMandalaSMat, "dgCMatrix")
testMandalaSMat <- drop0(testMandalaSMat)
dim(testMandalaSMat)
length(testMandalaSMat@x)/length(testMandalaSMat)
```

```{r}
image(matrix(as.numeric(t(testMandalaArray)), nrow = nrow(mandalaImageArrays[[1]]), byrow=F), axes=FALSE, col = gray(0:255/255))
```

## SVD

Represent by topics:

```{r}
matRepresentation <- 
  lsaSVDObj %>% 
  LSAMonNormalizeMatrixProduct(normalizeLeftQ = FALSE) %>% 
  LSAMonRepresentByTopics(query = testMandalaSMat) %>% 
  LSAMonTakeValue
```

```{r}
lsCoeff <- matRepresentation@x
barplot(lsCoeff)
```

```{r}
vecRepresentation = lsCoeff %*% svdH
dim(vecRepresentation)
summary(vecRepresentation@x)
```


```{r, message=FALSE}
vecRepresentation <- SMRApplyTermWeightFunctions(docTermMat = vecRepresentation, normalizerFunction = "AbsMax")
```


```{r, message=FALSE}
par( mfrow = c(1,2), mai = c(0,0,0,0))
image(matrix((1-vecRepresentation[1,]), nrow = nrow(mandalaImageArrays[[1]]), byrow=F), axes=FALSE, col = gray(0:255/255))
text( 0.2, 0, "approximated", cex = 1.4, col = 2, pos = c(3,4))
image(1-t(py$testMandalaArray), axes = FALSE, col = gray( 0:255 / 255 ))
text( 0.2, 0, "original", cex = 1.4, col = 2, pos = c(3,4))
```
## NNMF

Represent by topics:

```{r}
matRepresentation <- 
  lsaNNMFObj %>% 
  LSAMonNormalizeMatrixProduct(normalizeLeftQ = FALSE) %>% 
  LSAMonRepresentByTopics(query = testMandalaSMat, method = "Algebraic" ) %>% 
  LSAMonTakeValue
```

```{r}
lsCoeff <- matRepresentation@x
barplot(lsCoeff)
```

```{r}
vecRepresentation = lsCoeff %*% nnmfH
dim(vecRepresentation)
summary(vecRepresentation@x)
```


```{r, message=FALSE}
vecRepresentation <- SMRApplyTermWeightFunctions(docTermMat = vecRepresentation, normalizerFunction = "AbsMax")
summary(vecRepresentation@x)
vecRepresentation@x[vecRepresentation@x < 0] <- 0
```


```{r, message=FALSE}
par( mfrow = c(1,2), mai = c(0,0,0,0))
image(matrix((1-vecRepresentation[1,]), nrow = nrow(mandalaImageArrays[[1]]), byrow=F), axes=FALSE, col = gray(0:255/255))
text( 0.2, 0, "approximated", cex = 1.4, col = 2, pos = c(3,4))
image(1-t(py$testMandalaArray), axes = FALSE, col = gray( 0:255 / 255 ))
text( 0.2, 0, "original", cex = 1.4, col = 2, pos = c(3,4))
```