---
title: "QRMon anomalies detection"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: spacelab
    source_code: embed
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(magrittr)
library(ggplot2)
library(QRMon)
library(OutlierIdentifiers)

dfFocusAppleData <- read.csv( file = "../data/dfAppleMobilityLongForm.csv")

dfFocusAppleData <- 
  dfFocusAppleData %>% 
  dplyr::mutate( Regressor = as.numeric(as.POSIXct(DateObject)) )

```


Sidebar {.sidebar}
=====================================

```{r}
selectInput( inputId = "dataSpec", label = "Data:", choices = c( "Driving", "Transporation", "Walking" ), selected = "Driving" )

checkboxInput( inputId = "dateAxisQ", label = "Date axis:", value = FALSE )

checkboxInput( inputId = "joinedPlotQ", label = "Joined plot:", value = FALSE )

hr()

textInput( inputId = "probs", label = "Probabilities:", value = "0.1, 0.5, 0.9")

sliderInput( inputId = "df", label = "Degrees of freedom:", min = 1, max = 120, step = 1, value = 12 )

hr()

checkboxInput( inputId = "relativeErrorsQ", label = "Use relative errors:", value = FALSE )

numericInput( inputId = "outlierThreshold", label = "Outliers threshold:", min = 0, max = 100, step = 0.05, value = 1 )

selectInput( inputId = "outlierIdentifier", 
             label = "Outlier identifier:", 
             choices = c( "Hampel", "SPLUSQuartile", "Quartile"), 
             selected = "Hampel" )

```

Data and fit
=====================================

```{r}

lsGroupColors <- c( "Data" = "gray60", 
                    "VarianceDiff" = "gray60", 
                    "Outliers" = "red", 
                    "VarianceOutliers" = "red", 
                    "Anomalies" = "red" )

dataset <- reactive({
  
  res <- 
    dfFocusAppleData %>% 
    dplyr::filter( grepl( tolower(input$dataSpec), transportation_type) ) %>% 
    dplyr::select( Regressor, Value )
  
  QRMonUnit( setNames( res,  c("Regressor", "Value") ) ) %>% QRMonTakeData
})

outlierIdentifierParameters <- reactive({
  list( 
    "Hampel" = HampelIdentifierParameters, 
    "SPLUSQuartile" = SPLUSQuartileIdentifierParameters, 
    "Quartile" = QuartileIdentifierParameters
  )[[input$outlierIdentifier]]
})  

pointOutlierIdentifier <- reactive({
  function(x) TopOutlierIdentifier(data = x, identifier = outlierIdentifierParameters()) 
})

varianceOutlierIdentifier <- reactive({
  function(x) OutlierIdentifier(data = x, identifier = outlierIdentifierParameters()) 
})

lsProbs <- reactive({
  res <- strsplit( input$probs, ",|\\s")[[1]]
  as.numeric( res[nchar(res) > 0 ] )
})
  
qrObj <- reactive({

  QRMonUnit( dataset() ) %>%
  QRMonQuantileRegression( df = input$df, probabilities = lsProbs() )
})
```

### Summary

```{r}
renderPrint({
  cat( paste0("Number of points: ", nrow(dataset()), "\n"));
  summary(qrObj() %>% QRMonTakeData)
})
```


### Data and regression quantiles

```{r}
plotQR <- reactive({
  qrObj() %>% 
    QRMonPlot( echoQ = FALSE, 
               datePlotQ = input$dateAxisQ, 
               dateOrigin = "1900-01-01", 
               dataPointsColor = if( input$joinedPlotQ ) { NULL } else { "gray60" },
               dataLineColor = if( input$joinedPlotQ ) { "gray60" } else { NULL } ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  print(plotQR())
})
```

### Errors

```{r}
renderPlot( expr = {
  print(qrObj() %>% QRMonErrorsPlot(relativeErrorsQ = input$relativeErrorsQ, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ))
})
```

Point anomalies
=====================================

### Outliers by regression quantiles

```{r}
renderPlot( expr = {
  
  p <- 
    qrObj() %>% 
    QRMonOutliers %>% 
    QRMonOutliersPlot( 
      echoQ = FALSE, 
      plotRegressionCurvesQ = TRUE, 
      datePlotQ = input$dateAxisQ, 
      dateOrigin = "1900-01-01",
      dataPointsColor = if( input$joinedPlotQ ) { NULL } else { "gray60" },
      dataLineColor = if( input$joinedPlotQ ) { "gray60" } else { NULL } ) %>% 
    QRMonTakeValue
  
  print(p)
  
})
```

### Anomalies by residuals using threshold

```{r}
dfPointOutliers <- reactive({
  qrObj() %>% 
    QRMonFindAnomaliesByResiduals( threshold = input$outlierThreshold, outlierIdentifier = NULL, relativeErrorsQ = input$relativeErrorsQ ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  
  if( is.null(dfPointOutliers()) || nrow(dfPointOutliers()) == 0 ) {
    dfPlotData <- cbind( Type = "Data", dataset() )
  } else {
    dfPlotData <- rbind( cbind( Type = "Data", dataset() ), 
                         cbind( Type = "Outliers", dfPointOutliers() ) )
  }
  
  ggplot( dfPlotData ) + 
    geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                     y = Value, 
                     color = Type ) ) +
    scale_colour_manual( values = lsGroupColors ) +
    xlab("Regressor")
                
})
```

### Anomalies by residuals using outlier identifier

```{r}
dfPointOutliersByFunc <- reactive({
  qrObj() %>% 
    QRMonFindAnomaliesByResiduals( threshold = NULL, outlierIdentifier = pointOutlierIdentifier(), relativeErrorsQ = input$relativeErrorsQ ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
   
  if( is.null(dfPointOutliers()) || nrow(dfPointOutliersByFunc()) == 0 ) {
    dfPlotData <- cbind( Type = "Data", dataset() )
  } else {
    dfPlotData <- rbind( cbind( Type = "Data", dataset() ), 
                         cbind( Type = "Outliers", dfPointOutliersByFunc() ) )
  }
  
  ggplot( dfPlotData ) + 
    geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                     y = Value, 
                     color = Type ) ) +
    scale_colour_manual( values = lsGroupColors ) +
    xlab("Regressor")
  
})
```


Variance anomalies
=====================================

### Regression quantiles for variance evaluation

```{r}
renderPlot( expr = {
  
  print(
    QRMonUnit( qrObj() %>% QRMonTakeData ) %>% 
      QRMonQuantileRegression( df = input$df, probabilities = c( min(lsProbs()), max(lsProbs()) ) ) %>% 
      QRMonPlot( echoQ = FALSE, 
                 datePlotQ = input$dateAxisQ, 
                 dateOrigin = "1900-01-01",
                 dataPointsColor = if( input$joinedPlotQ ) { NULL } else { "gray60" },
                 dataLineColor = if( input$joinedPlotQ ) { "gray60" } else { NULL } ) %>% 
      QRMonTakeValue 
  )
  
})
```

### Conditional variance

```{r}
renderPlot( expr = {
  
  lsRes <- qrObj() %>% QRMonPredict( ) %>% QRMonTakeValue
  dfRes <- dplyr::bind_rows( lsRes, .id = "Prob")
  
  maxPPos <- which.max( as.numeric(names(lsRes)) )
  minPPos <- which.min( as.numeric(names(lsRes)) )

  dfDiffs <- data.frame( Regressor = lsRes[[maxPPos]]["Regressor"], Value = abs(lsRes[[maxPPos]]["Value"] - lsRes[[minPPos]]["Value"] ) )

  lsOutlierPos <- varianceOutlierIdentifier()( dfDiffs$Value )
  
  if(length(lsOutlierPos) == 0 || sum(lsOutlierPos) == 0 ) {
    dfPlotData <- cbind( Type = "Data", dfDiffs )
  } else {
    dfPlotData <-  rbind( cbind( Type = "VarianceDiff", dfDiffs ), 
                          cbind( Type = "Outliers", dfDiffs[lsOutlierPos, ,drop=F] ) )
  }
  
  ggplot( dfPlotData ) + 
    geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                     y = Value,  
                     color = Type)
    ) +
    scale_colour_manual( values = lsGroupColors ) +
    xlab("Regressor")
    
})
```

### Variance outliers

```{r}
dfVarOutliers <- reactive({
  qrObj() %>% 
    QRMonFindVarianceAnomalies( outlierIdentifier = varianceOutlierIdentifier(), positionsQ = FALSE ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  
  if( nrow(dfVarOutliers()) == 0 ) {
    dfPlotData <- cbind( Type = "Data", dataset() )
  } else {
    dfPlotData <-  rbind( cbind( Type = "Data", dataset() ), 
                          cbind( Type = "VarianceOutliers", dfVarOutliers() ) )
  }
  
  ggplot( dfPlotData ) + 
    geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                     y = Value, 
                     color = Type ) ) +
    scale_colour_manual( values = lsGroupColors ) +
    xlab("Regressor")
  
})
```

References
====================================

#### In brief

The quantile regression workflows are made with the R package [`QRMon-R`](https://github.com/antononcube/QRMon-R),
which is based on [`quantreg`](https://cran.r-project.org/web/packages/quantreg/index.html).

#### Dashboards

[AAd1] Anton Antonov,
[GNNMon time series anomalies detection](https://antononcube.shinyapps.io/GNNMonTimeSeriesAnomaliesDetection/),
(2020),
[shinyapps.io by RStudio](https://www.shinyapps.io).

#### Articles, books

[RK1] Roger Koenker, 
[Quantile Regression](https://books.google.com/books/about/Quantile_Regression.html?id=hdkt7V4NXsgC), 
Cambridge University Press, 2005.

[RK2] Roger Koenker,
["Quantile Regression in R: a vignette"](https://cran.r-project.org/web/packages/quantreg/vignettes/rq.pdf),
(2006),
[CRAN](https://cran.r-project.org/).

[AA1] Anton Antonov,
["A monad for Quantile Regression workflows"](https://github.com/antononcube/MathematicaForPrediction/blob/master/MarkdownDocuments/A-monad-for-Quantile-Regression-workflows.md),
(2018),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

#### Packages

[RKp1] Roger Koenker,
[`quantreg`](https://cran.r-project.org/web/packages/quantreg/index.html),
[CRAN](https://cran.r-project.org/).

[AAp1] Anton Antonov,
[Quantile Regression Mathematica package](https://github.com/antononcube/MathematicaForPrediction/blob/master/QuantileRegression.m),
(2014),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

[AAp2] Anton Antonov,
[Monadic Quantile Regression Mathematica package](https://github.com/antononcube/MathematicaForPrediction/blob/master/MonadicProgramming/MonadicQuantileRegression.m),
(2018),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

[AAp3] Anton Antonov,
[`QuantileRegression`](https://resources.wolframcloud.com/FunctionRepository/resources/QuantileRegression),
(2019),
[Wolfram Function Repository](https://resources.wolframcloud.com/FunctionRepository/resources/QuantileRegression).

#### Repositories

[AAr1] Anton Antonov,
[QRMon-R](https://github.com/antononcube/QRMon-R),
(2019),
[GitHub/antononcube](https://github.com/antononcube).

[AAr2] Anton Antonov,
[DSL::English::QuantileRegressionWorkflows in Raku](https://github.com/antononcube/Raku-DSL-English-QuantileRegressionWorkflows),
(2020),
[GitHub/antononcube](https://github.com/antononcube).

#### Videos

[AAv1] Anton Antonov,
["Anomalies, Breaks, and Outliers Detection in Time Series"](https://www.youtube.com/watch?v=h_fLb6YU87c),
Wolfram Research Technology 2019 talk,
[Wolfram Research Inc. channel at YouTube](https://www.youtube.com/channel/UCJekgf6k62CQHdENWf2NgAQ).

[AAv2] Anton Antonov,
["Boston useR! QuantileRegression Workflows 2019-04-18"](https://www.youtube.com/watch?v=a_Dk25xarvE),
(2019),
[Anton Antonov at YouTube](https://www.youtube.com/channel/UC5qMPIsJeztfARXWdIw3Xzw).

[AAv3] Anton Antonov,
["useR! 2020: How to simplify Machine Learning workflows specifications"](https://www.youtube.com/watch?v=b9Uu7gRF5KY),
(2020),
[R Consortium at YouTube](https://www.youtube.com/channel/UC_R5smHVXRYGhZYDJsnXTwg).
