---
title: "Scaffolding examples"
author: Anton Antonov
date: 2021-02-04
output: html_notebook
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r setup}
library(shiny)
library(tidyverse)
library(magrittr)
library(QRMon)
library(OutlierIdentifiers)
library(ROCFunctions)
```

# Time series data

## Ingestion

```{r}
dfTimeSeriesWithOutliers <- read.csv( file = "~/MathFiles/Presentations/Champaign-Urbana-DataScienceUserGroup-Feb-2021/output/dsTimeSeriesWithOutliers.csv")
```

```{r}
# dfTimeSeriesWithOutliers <- 
#   dfTimeSeriesWithOutliers %>% 
#   dplyr::mutate( Type = ifelse( BottomOutlier == "True", "BottomOutlier", "None") ) %>% 
#   dplyr::mutate( Type = ifelse( TopOutlier == "True", "TopOutlier", Type) )
```

```{r}
summary(as.data.frame(unclass(dfTimeSeriesWithOutliers), stringsAsFactors = T), maxsum = 12)
```

## Plots

```{r, fig.width = 12, fig.height = 8, fig.fullwidth=TRUE, message=FALSE}
ggplot(dfTimeSeriesWithOutliers) +
  geom_point( aes( x = as.POSIXct(Regressor, origin = "1900-01-01"), y = Value, color = Type ) ) +
  xlab("Regressor") +
  facet_wrap( ~ID, ncol = 2, scales = "free" )
```

## Focus time series

Here we pick one of the time series to do the computations below:

```{r}
dfFocusTimeSeries <- 
  dfTimeSeriesWithOutliers %>% 
  dplyr::filter( ID == "hvps-kids-lsro" ) %>% 
  dplyr::mutate( ID = NULL )
dfFocusTimeSeries
```

-----

# Find anomalies

## QRMon object

Here we do the following steps:

1. Create a `QRMon` object with the selected time series
2. Compute a fit at probability 0.5
3. Plot data and fit

```{r, fig.width = 8, fig.height = 2, fig.fullwidth=TRUE}
qrObj <- 
  QRMonUnit(dfFocusTimeSeries) %>% 
  QRMonQuantileRegression(df = 20, probabilities = c(0.5) ) %>% 
  QRMonPlot( datePlotQ = TRUE, dateOrigin = "1900-01-01")
```

## Batch testing of the anomalies finding algorithm

We want to see the performances of the anomalies finding algorithm (method) with different combinations of parameter values.

Here we make a data frame with different parameter values, each row is a set of parameters:

```{r}
dfParams <- 
    expand.grid( c( NA, 0.5, 1, 3, 4, 5, 10),
                 c( "HampelIdentifierParameters", "SPLUSQuartileIdentifierParameters", "QuartileIdentifierParameters" ), 
                 c( TRUE, FALSE), 
                 stringsAsFactors = FALSE )
dfParams <- setNames( dfParams, c("Threshold", "OutlierIdentifier", "RelativeErrorsQ") )
dfParams <- 
  dfParams %>% 
  dplyr::mutate( OutlierIdentifier = ifelse( !is.na(Threshold), NA, OutlierIdentifier ) ) %>% 
  unique() %>% 
  dplyr::mutate( ID = dplyr::row_number() ) %>% 
  dplyr::select( ID, names(dfParams))
dfParams
```

Here we make a dictionary that maps parameters combinations into labels:

```{r}
aIDToLabel <- purrr::map( split(dfParams,dfParams$ID), function(x) do.call( paste, as.list(x[,-1]) ) )
```

Find anomalies for each parameter set:

```{r}
lsFoundOutliers <- 
  purrr::map( split(dfParams, dfParams$ID), function(dfPars) {
      cbind(
        ID = dfPars$ID,
        qrObj %>% 
          QRMonFindAnomaliesByResiduals( threshold = if( is.na(dfPars$Threshold) ) { NULL } else { dfPars$Threshold }, 
                                         outlierIdentifier = if( !is.function(dfPars$OutlierIdentifier) ) { NULL } else { parse( text = dfPars$OutlierIdentifier ) },
                                         relativeErrorsQ = dfPars$RelativeErrorsQ ) %>% 
          QRMonTakeValue
      )
  })
length(lsFoundOutliers)
```

Here is an example of one of the results from the batch run above:

```{r}
lsFoundOutliers[[2]]
```

Compare with the known anomalies:

```{r}
dfFoundOutliersCounts <- 
  purrr::map_df( lsFoundOutliers, function(dfFound) {
  res <- 
    dfFocusTimeSeries %>% 
    dplyr::mutate( ID = dfFound$ID[[1]], Type = gsub( ".*Outlier", "Outlier", Type ) ) %>% 
    dplyr::left_join( dfFound %>% dplyr::mutate( ID = NULL, FoundType = "Outlier"), by = c("Regressor", "Value") ) %>% 
    dplyr::mutate( FoundType = ifelse( is.na(FoundType), "Normal", FoundType ) ) %>%
    dplyr::count( ID, Type, FoundType, name = "Count" ) %>% 
    dplyr::inner_join( 
      dplyr::tribble(
        ~Type,     ~FoundType, ~ROCType,
        "Normal",  "Normal",   "TrueNegative",
        "Outlier", "Outlier",  "TruePositive",
        "Outlier", "Normal",   "FalseNegative",
        "Normal",  "Outlier",  "FalsePositive"
      ), by = c("Type", "FoundType"))
})
dfFoundOutliersCounts
```

Convert to wide form (more convenient for the computations below):

```{r}
dfFoundOutliersCountsWide <- tidyr::pivot_wider( data = dfFoundOutliersCounts, id_cols = "ID", names_from = "ROCType", values_from = "Count" )
dfFoundOutliersCountsWide[ is.na(dfFoundOutliersCountsWide) ] <- 0
dfFoundOutliersCountsWide
```

Compute 
[Receiver Operating Characteristic (ROC)](https://en.wikipedia.org/wiki/Receiver_operating_characteristic) functions 
using the R-package 
[ROCFunctions](https://github.com/antononcube/R-packages/tree/master/ROCFunctions):

```{r}
dfROCs <- ROCFunctions::ComputeROCFunctions(dfFoundOutliersCountsWide, c("FPR", "TPR", "PPV"))
dfROCs
```

Plot False Positive Rate (FPR) vs True Positive Rate (TPR):

```{r}
ggplot(dfROCs) +
  geom_point( aes( x = FPR, y = TPR ) ) +
  geom_label( aes( x = FPR, y = TPR, label = aIDToLabel[ID]), nudge_x = 0.04, nudge_y = -0.04) +
  xlim(-0.2, 1.2) + ylim(-0.2, 1.2)
```

# Custom made measure (and an interactive interface)

We might want to do a weighted sum of True Positive Rate (TPR) and Positive Predictive Value (PPV):

$$\frac{1}{5} \frac{TP}{TP+FN} + \frac{4}{5} \frac{TP}{TP+FP}$$
(TPR is the same Recall.)

```{r}
app <- 
  shinyApp(
    ui = fluidPage(
      sliderInput( inputId = "theta", label = "theta:", min = 0, max = 1, value = 0.5, step = 0.01),
      plotOutput("plot")
    ),
    server = function(input, output) {
      output$plot <- renderPlot(
       ggplot(dfROCs) +
         geom_point( aes( x = ID, y = input$theta * TPR + (1 - input$theta) * PPV ) ) +
         geom_label( aes( x = ID, y = input$theta * TPR + (1 - input$theta) * PPV, label = aIDToLabel[ID]), nudge_x = 0.04, nudge_y = -0.04) +
         xlab("ID") + ylab("Score") +
         ylim(0,1.2)
      )
    }
  )

runApp(app)
```


